Boosting Higgs Pair Production in the $b\bar{b}b\bar{b}$ Final State with Multivariate Techniques

### 2015-12-04 paper/sec-analysis.tex
more work on paper

In section we describe our analysis strategy, and in particular
the classification of each individual event into
different categories according to its topology.
%
First of all, we discuss the settings and nomenclature that
we will use for jet clustering, and then we move to discuss how $b$-tagging is simulated.
%
Then we discuss the categorisation of events between different
topologies, and how we prioritise among them.
%
Finally, we motivate our choice of analysis cuts by comparing signal and background
for some representative kinematic distributions.

@@

At the end of the  parton shower, final state particles
are clustered using the
jet reconstruction algorithms
are obtained from the
{\tt FastJet} program~\cite{Cacciari:2011ma,Cacciari:2005hq},
{\tt v3.1.0}.
%
In this study we will use different types of jet definitions,
which we enumerate now:
%

\begin{itemize}
\item {\it Small-$R$ jets}.

  These are jets  reconstructed with the
  anti-$k_T$ clustering algorithm~\cite{Cacciari:2008gp} with $R=0.4$ radius.
  %
  When reconstructing small-$R$ jets, we keep only
  those jets with transverse momentum $p_T \ge 40$~GeV
  and pseudo-rapidity $|\eta|<2.5$, within the central rapidity
  acceptance of ATLAS and CMS.
%
  Jets that fail to satisfy these acceptance
  requirements are discarded.
  %
  The restriction to the central rapidity region is a
  requisite for $b$-tagging.

@@

  These jets are also constructed with the
  anti-$k_T$ clustering algorithm, now using a $R=1.0$ radius.
  %
  In order to be accepted, these large-$R$ jets should
  satisfy  $p_T \ge 200$~GeV and lie in a pseudo-rapidity range of
  $|\eta|<2.0$.
  %
  If these conditions are not satisfied, the jet is discarded.
  %
  The more restrictive range  in pseudo-rapidity
  as compared to the small-$R$ jets
  is motivated by mimicking the  experimental requirements
  in ATLAS and CMS
  related to the track-jet-based calibration~\cite{Aad:2014bia,ATLAS:2012kla}.

  In addition to the basic $p_T$ and $\eta$
  acceptance requirements, large-$R$ jets should also
  satisfy the  BDRS mass-drop tagger~\cite{Butterworth:2008iy}
  conditions, required
  to enhance the discrimination of the signal over the background
  events.
  %
  For the BDRS mass-drop tagger, we use the {\tt FastJet} default
  parameters of  $\mu = 0.67$ and $y_{\textrm{cut}}= 0.09$.
  %
  Before the mass-drop tagger is applied, the large-$R$ jet
  constituents are reclustered with the Cambridge/Aachen
  algorithm~\cite{Dokshitzer:1997in,Wobisch:1998wt}.

@@

  Once we have reconstructed a suitable large-$R$ jet, important
  information can be obtained by looking at the kinematics of
  the small-$R$ subjets.
  %
  These subjets will be obtained by reclustering the constituents
  of the large-$R$ jet, always with the  anti-$k_T$ algorithm,
  but this time with a smaller radius parameter $R=0.3$.
 %
  These small-$R$ subjets will be the building blocks of
  $b$-tagging in the boosted category.
  %
  Small-$R$ subjets are required to have $p_T > 50$~GeV and $|\eta|<2.5$,
  else they are discarded.
  %
\end{itemize}

We have
also explored the possible improvements
in the analysis from using variable-$R$ jets~\cite{Krohn:2009zg}, but found that these made
a small difference as compared to fixed-$R$ jets.
%
This is not unexpected, since variable-$R$ jets are more suitable when
the degree of boost of the final state being reconstructed can span
a wide range, such as in heavy resonance searches~\cite{Aad:2015fna}.

\subsection{$b$-tagging}
\label{sec:btagging}

Given the final state that needs to be reconstructed, the
optimisation of the $b$-tagging capabilities of the
LHC experiments is an essential requirement to improve
the prospects for the observation of Higgs pair production in this channel.
%
The $b$-tagging strategy used in this feasibility study is inspired
in the current and projected ATLAS settings~\cite{Aad:2013gja},
though
it is not too different from
 the corresponding CMS strategy~\cite{Khachatryan:2011wq,Chatrchyan:2012jua}.
%
For each type of jet defined above, a different
$b$-tagging strategy is used:

@@

  These are $b$-tagged as follows.
  %
  If a small-$R$ jet has at least one $b$-quark among their constituents,
  it will be tagged as a $b$-jet with probability $f_b$.
  %
  In order to be considered in the $b$-tagging algorithm,
   $b$-quarks should satisfy $p_T \ge 15$ GeV.
  %
  If no $b$-quarks are found among the constituents
  of this jet, it can be still be tagged as a $b$-jet with
  a mistag rate of $f_l$, unless a charm quark is found,
  and in this case the mistag rate is $f_c$.
  %
  Only jets which contain at least one constituent
  with $p_T \ge 15$ GeV can lead to fake $b$-tags.
  %
  The probability of tagging a jet is not modified
  if more than one $b$-quark is found among the jet constituents.

  We only attempt to $b$-tag the four (two) hardest small-$R$ jets
  in the resolved (intermediate) category.
  %
  We find that
  attempting to $b$-tag all
  small-$R$ jets that satisfy the acceptance cuts worsens the performance
  due to combinatorics.

@@

    Large-$R$ jets are $b$-tagged by
    ghost-associating anti-$k_T$ $R=0.3$
    subjets to the original large-$R$
    jets~\cite{Cacciari:2007fd,Aad:2013gja,
      ATLAS-CONF-2014-004}.\footnote{The same
    method was used in the recent ATLAS $hh\to 4b$ search~\cite{Aad:2015uka}.}
    %
    In particular,
    a large-$R$ jet is considered  double-$b$-tagged if both
    the leading and subleading subjets (where the ordering
    is done in the subjet $p_T$) are both individually $b$-tagged.

    As in the case
    of small-$R$ jets, in
    order to be considered in the $b$-tagging algorithm,
   $b$-quarks should satisfy $p_T \ge 15$ GeV.
    %
    Note that we only attempt to $b$-tag the two leading subjets:
    trying to $b$-tag all subjets above the $p_T$ cut degrades
    the signal significance due to the increase in combinatorics.

    Concerning the $b$-tagging probabilities and the
    light jet  and charm jet mistag probabilities, there are taken
    the same as for small-$R$ jets.
    %
    Therefore, a large-$R$ jet where the two leading
    subjets have at least one $b$-quark will be tagged
    with probability $f_b^2$, if only one of the two leading
    subjets has a $b$-quark then the tagging probability is
    $2f_bf_l$, and if none of the two have $b$-quarks
    as constituents, the mistag rate will be
    $f_l^2$.

@@

Concerning the values of the $b$-tag probability $f_b$ and
the $b$-mistag probability of light and charm jets $f_l$ and  $f_c$, 
in this work we adopt as baseline the values $f_b=0.8$, $f_l=0.01$
and  $f_c=0.1$.

@@

The basic idea of our analysis strategy is similar to that in Ref.~\cite{Gouzevitch:2013qca}:
rather than concentrating only a specific event topology, such as boosted
or resolved, we aim to consistently combine the information from
the three possible event topologies: boosted, intermediate and
resolved.
%
However, an important difference compared to~\cite{Gouzevitch:2013qca}
is that here each category is optimised separately, and then
a exclusive categorisation is defined starting from the category
with the highest signal significance.
%
An advantage of our approach is that the analysis strategy would be unchanged
if the underlying production model of Higgs pairs is modified, for example in the case of
BSM dynamics, which can substantially increase the degree of boost of the final state.

>>>>

In section we describe our analysis strategy, and in particular
the classification of each individual event into
different categories according to its topology.
%
First of all, we discuss the settings and nomenclature
for jet clustering, and the strategy for $b$-tagging.
%
Then we discuss the categorisation of events between different
topologies and how to prioritise among them.
%
Finally, we motivate our choice of analysis cuts by comparing signal and background
for representative kinematic distributions.

@@

After the  parton shower, final state particles
are clustered using the
jet reconstruction algorithms
obtained from
{\tt FastJet}~\cite{Cacciari:2011ma,Cacciari:2005hq},
{\tt v3.1.0}.
%
We will use different types of jet definitions:
\begin{itemize}
\item {\it Small-$R$ jets}.

  These are jets  reconstructed with the
  anti-$k_T$ clustering algorithm~\cite{Cacciari:2008gp} with $R=0.4$ radius.
  %
  These small-$R$ jets are required
  to have transverse momentum $p_T \ge 40$~GeV
  and pseudo-rapidity $|\eta|<2.5$, within the central 
  acceptance of ATLAS and CMS, the region
  where $b$-tagging is possible.

@@

  These jets are also constructed with the
  anti-$k_T$ clustering algorithm, now using a $R=1.0$ radius.
  %
  Large-$R$ jets are required to have
  $p_T \ge 200$~GeV and lie in a pseudo-rapidity region of
  $|\eta|<2.0$.
  %
  The more restrictive range  in pseudo-rapidity
  as compared to the small-$R$ jets
  is motivated by mimicking the  experimental requirements
  in ATLAS and CMS
  related to the track-jet based calibration~\cite{Aad:2014bia,ATLAS:2012kla}.

  In addition to the basic $p_T$ and $\eta$
  acceptance requirements, large-$R$ jets should also
  satisfy the  BDRS mass-drop tagger (MDT)(~\cite{Butterworth:2008iy}
  conditions.
  %
  We use the {\tt FastJet} default
  parameters of  $\mu = 0.67$ and $y_{\textrm{cut}}= 0.09$.
  %
  Before applying the MDT, the large-$R$ jet
  constituents are reclustered with the Cambridge/Aachen (C/A)
  algorithm~\cite{Dokshitzer:1997in,Wobisch:1998wt}
  with $R=1.2$.

@@

  These subjets are constructed by reclustering the constituents
  of a large-$R$ jet, again with  anti-$k_T$ algorithm,
  but this time with a smaller radius parameter, namely
  $R=0.3$.
 %
  These small-$R$ subjets will be the main input for
  $b$-tagging in the boosted category.
  %
  They are required to satisfy $p_T > 50$~GeV and $|\eta|<2.5$.
  %
\end{itemize}

In addition, we have
also explored the possible improvements
in the analysis from the use
of variable-$R$ jets~\cite{Krohn:2009zg}, but found that these made
a small difference as compared to fixed-$R$ jets.
%
This is not unexpected, since variable-$R$ jets are more suitable when
the degree of boost of the final state being reconstructed can span
a wide range~\cite{Cacciari:2008gd}, such as in heavy resonance searches~\cite{Aad:2015fna},
where where the degree of boost of the final state
is {\it a priori} unknown.

\subsection{Tagging of $b$-jets}
\label{sec:btagging}

It is clear that
optimisation of the $b$-tagging capabilities of the
LHC experiments is an essential ingredient to improve
the prospects for the observation of Higgs pair production in the
$b\bar{b}b\bar{b}$ final state.
%
In this work, we adopt
a $b$-tagging strategy along the lines
of current ATLAS performance~\cite{Aad:2013gja,Aad:2015ydr},
though differences with respect to
the corresponding CMS strategy~\cite{Khachatryan:2011wq,Chatrchyan:2012jua}
should not affect qualitatively our results.
%
For each type of jet definition used, a different
$b$-tagging strategy is adopted:

@@

  %
  If a small-$R$ jet has at least one $b$-quark among their constituents,
  it will be tagged as a $b$-jet with probability $f_b$.
  %
  In order to be considered in the $b$-tagging algorithm,
  $b$-quarks inside the small-$R$ jet
  should satisfy $p_T \ge 15$ GeV.
  %
  The probability of tagging a jet is not modified
  if more than one $b$-quark is found among the jet constituents.

  If no $b$-quarks are found among the constituents
  of this jet, it can be still be tagged as a $b$-jet with
  a mistag rate of $f_l$, unless a charm quark is present instead,
  and in this case the mistag rate is $f_c$.
  %
  Only jets which contain at least one (light or charm)
  constituent
  with $p_T \ge 15$ GeV will lead to fake $b$-tags.

  We only attempt to $b$-tag only the four (two) hardest small-$R$ jets
  in the resolved (intermediate) category.
  %
  Trying to $b$-tag all the
  small-$R$ jets that satisfy the acceptance cuts worsens the
  overall performance
  due to combinatorics.

@@

    Large-$R$ jets are $b$-tagged by
    ghost-associating anti-$k_T$ $R=0.3$
    subjets to the original large-$R$
    jets~\cite{Cacciari:2007fd,Aad:2013gja,
      ATLAS-CONF-2014-004,Aad:2015uka}.
    %
    A large-$R$ jet will considered $b$-tagged if both
    the leading and subleading AK03 subjets, where the ordering
    is done in the subjet $p_T$, are both individually $b$-tagged,
    with the same criteria as the small-$R$ jets.

    As in the case
    of small-$R$ jets, we only attempt to $b$-tag the two leading subjets,
    else one finds a degradation of the
    signal significance due to combinatorics.

    %
    Therefore, a large-$R$ jet where the two leading
    subjets have at least one $b$-quark will be tagged
    with probability $f_b^2$, if only one of the two leading
    subjets has a $b$-quark, then the tagging probability is
    $2f_bf_l$, and if none of the two subjets include $b$-quarks,
    the mistag rate will be $f_l^2$.

@@

Concerning the values of the $b$-tag probability, $f_b$, and
the $b$-mistag probability of light and charm jets, $f_l$ and  $f_c$
respectively,
we used the values $f_b=0.8$, $f_l=0.01$
and  $f_c=0.1$.

@@

The basic idea of present analysis is similar to the
scale-invariant resonance tagging introduced
in Ref.~\cite{Gouzevitch:2013qca}.
%
rather than restricting to a specific event topology, such as boosted
or resolved, we aim to consistently combine the information from
the three possible event topologies: boosted, intermediate and
resolved.
%
However, an important difference compared to~\cite{Gouzevitch:2013qca}
is that here each category is optimised separately, and then
a exclusive categorisation is defined starting from the category
with the highest signal significance.
%
An advantage of our approach is that the analysis strategy would be unchanged
if the underlying production model of Higgs pairs is modified, for example in the case of
BSM dynamics, which can substantially increase the degree of boost of the final state.

### 2015-12-04 paper/sec-pileup.tex
more work on paper

To study the impact of PU in our analysis,
Minimum Bias (MB) events,
including Multiple Parton Interactions (MPI), have been generated
with {\tt Pythia8}, and
superimposed to the signal
and background samples described in Sect.~\ref{mcgeneration}.
%
We have explored two scenarios for the amount of PU expected
at the HL-LHC, one with a mean number of
PU vertices per bunch crossing of $\la n_{\rm PU}\ra=80$, and another
with $\la n_{\rm PU}\ra=150$.
%
In order to subtract the PU, a number of techniques
have been developed
recently~\cite{Cacciari:2009dp,TheATLAScollaboration:2013pia,Butterworth:2008iy,Cacciari:2007fd,Krohn:2009th,Krohn:2013lba,Ellis:2009me,Bertolini:2014bba,Cacciari:2014gra,Cacciari:2014jta,Berta:2014eza,Larkoski:2014wba}.
%
In this work, PU  will be subtracted by means
of the the {\tt SoftKiller} (SK)
method~\cite{Cacciari:2014gra}, as implemented in {\tt FastJet}.
%

>>>>

To study the impact of PU in our analysis,
Minimum Bias (MB) events,
including Multiple Parton Interactions (MPI), have been generated
with {\tt Pythia8}, and
superimposed to the signal
and background samples described in Sect.~\ref{mcgeneration}.
%
We have explored two scenarios for the amount of PU expected
at the HL-LHC, one with a mean number of
PU vertices per bunch crossing of $\la n_{\rm PU}\ra=80$, and another
with $\la n_{\rm PU}\ra=150$.
%
In order to subtract the PU, a number of techniques
have been developed
recently~\cite{Cacciari:2009dp,TheATLAScollaboration:2013pia,Butterworth:2008iy,Cacciari:2007fd,Krohn:2009th,Krohn:2013lba,Ellis:2009me,Bertolini:2014bba,Cacciari:2014gra,Cacciari:2014jta,Berta:2014eza,Larkoski:2014wba}.\footnote{
These techniques have also important applicability for the subtraction
of UE/MPI contamination for jet reconstruction
in heavy ion collisions~\cite{Cacciari:2010te}.
}
%
In this work, PU  will be subtracted by means
of the the {\tt SoftKiller} (SK)
method~\cite{Cacciari:2014gra}, as implemented in {\tt FastJet}.
%
